---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# forks

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/asshah4/forks/branch/main/graph/badge.svg)](https://codecov.io/gh/asshah4/forks?branch=main)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/asshah4/forks/workflows/R-CMD-check/badge.svg)](https://github.com/asshah4/forks/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/forks)](https://CRAN.R-project.org/package=forks)
<!-- badges: end -->

_Formulas with Operations Prescribed_ are described within the `{forks}` package, $/fɔrks/$, which helps to extend the base R `formula` class.

## Installation

You can install the released version of `forks` from [CRAN](https://CRAN.R-project.org) with:

```{r}
#install.packages("forks")
```

And the development version from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("asshah4/forks")
```

## Usage

The `forks` package is simple, and only has several major functions.

- `rx()`
- `set_rx_labels()`

It simply adds a layer of attributes to `formula` objects that can help simplify their usage in other packages, particularly in the field of epidemiology and causality-based modeling.

```{r}
library(rx)
# Create a simple formula
f <- mpg ~ wt + hp + cyl
new_f <- rx(f)
class(new_f)

# Identifies the roles of individual terms
term_tbl <- attributes(new_f)$roles

# These can also be listed out as labels if need be
table_to_list(term_tbl)
```

For other purposes, the individual terms may need to be labeled. For example, in the `murmur` package, exposures and fixed variables are treated differently.

```{r}
# This can be done manually
set_rx_labels(labels = list(
	X = "exposures",
	F = "fixed"
))

f <- mpg ~ X(wt) + hp + F(cyl)
new_f <- rx(f)

# Identifies the roles of individual terms
attributes(new_f)$roles %>%
	table_to_list()

# This can also be done through "themes" based on individual packages
reset_rx_labels()
set_rx_theme("murmur")
f <- mpg ~ X(wt) + hp + F(cyl)
new_f <- rx(f)
attributes(new_f)$roles %>%
	table_to_list()
```


