---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# axe

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/asshah4/axe/branch/main/graph/badge.svg)](https://codecov.io/gh/asshah4/axe?branch=main)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/asshah4/axe/workflows/R-CMD-check/badge.svg)](https://github.com/asshah4/axe/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/axe)](https://CRAN.R-project.org/package=axe)
<!-- badges: end -->

_Formulas with prescribed operations_ are described within the `{axe}` package, which helps to extend and vectorize the base R `formula` class. The purpose is to help __chop__ a formula-based object into underlying `term` objects, and __rebuild__ it as _prescribed formula_.

## Installation

You can install the released version of `axe` from [CRAN](https://CRAN.R-project.org) with:

```{r}
#install.packages("axe")
```

And the development version from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("asshah4/axe")
```

## Usage

The `axe` package is simple, and only has several major functions.
