---
title: "Development Thoughts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Development Thoughts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Statement of Need

The _raison d'étre_ of the `{forks}` package, pronounced __/fɔrks/__ ...

# Development

The `{forks}` package was developed as an extension of the `formula` class in base R. The `{vctrs}` package was used to extend the approach to formula handling.

## Considerations

These are some of the thoughts that led to the development of this package.

- The base `formula` class, as it applies to modeling and relationship descriptions
- Expansion of formulas into groups/lists of related formulas
- Fixed or protected roles of individual terms (e.g. outcomes, exposures, predictors, mediators, confounders)
- Customized or unprotected roles for individual terms (specific for individual packages)
- Operations that applies to individual terms and their distribution at the level of the data itself
- Operations that apply to an overall formula itself, affecting the relationships between terms
- Labels for individual terms (e.g. `mtcars$cyl ~ "Number of Cylinders"`)

## Vectors

As `{vctrs}` forms the base for this, a key concept is the idea of vector types and __coercion hierarchy__.

<center>

```{r, echo=FALSE, fig.align='center'}
DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = BT]
node [shape = rectangle, fontname = arial]
c [label = character, color = white]
t [label = <<b>TERMS</b>>, color = white, fontcolor = crimson]
f [label = <<b>FORMULA</b>>, color = white, fontcolor = crimson]
m [label = model, color = white]
fl [label = 'formula list', color = white]
ml [label = 'model list', color = white]

c -> t [color = rosybrown]
t -> f [color = rosybrown]
f -> fl [color = rosybrown]
f -> m [color = rosybrown]
fl -> ml [color = rosybrown]
m -> ml [color = rosybrown]

}")
```

</center>

In theory, the `character` class is considered enriched as a `term` class. In turn, a `term` class is considered enriched as a `formula` class. These enriched classes of <span style = "color:crimson">term</span> and <span style = "color:crimson">formula</span> could be implemented as vectors under this type of hierarchy, with emergent properties that develop as we ascend.

