---
title: "development"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, setup, warning=FALSE, echo = FALSE}
library(forks)
library(nomnoml)
library(DiagrammeR)
```

# Statement of Need

The _raison d'étre_ of the `{forks}` package, pronounced __/fɔrks/__ ...

# Development

The `{forks}` package was developed as an extension of the `formula` class in base R. The `{vctrs}` package was used to extend the approach to formula handling.

## Considerations

These are some of the thoughts that led to the development of this package.

- The base `formula` class, as it applies to modeling and relationship descriptions
- Expansion of formulas into groups/lists of related formulas
- Fixed or protected roles of individual terms (e.g. outcomes, exposures, predictors, mediators, confounders)
- Customized or unprotected roles for individual terms (specific for individual packages)
- Operations that applies to individual terms and their distribution at the level of the data itself
- Operations that apply to an overall formula itself, affecting the relationships between terms
- Labels for individual terms (e.g. `mtcars$cyl ~ "Number of Cylinders"`)

## Vectors

As `{vctrs}` forms the base for this, a key concept is the idea of vector types and __coercion hierarchy__.

<center>

```{r, echo=FALSE, fig.align='center'}
DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = BT]
node [shape = plaintext, fontname = arial]
edge [color = rosybrown]

# Terms
{
	rank=same
	ch [label = <<b>CHARACTER</b>>]
	ls [label = label]
	rl [label = role]
}
tx [label = <<b>TERM</b>>, fontcolor = crimson]
tl [label = <<b>TERM LIST</b>>, fontcolor = crimson]

# Formulas
fx [label = <<b>FORMULA</b>>, fontcolor = crimson]
pt [label = pattern]
fl [label = <<b>FORMULA LIST</b>>, fontcolor = crimson]

# Models
mx [label = <<b>MODEL</b>>]
ml [label = <<b>MODEL LIST</b>>, fontcolor = crimson]
df [label = 'data']

# Relationships
subgraph skeleton{
	ch -> tx
	tx -> ch 
	tx -> tl 
	tl -> tx 
	tl -> fx 
	fx -> tl 
	fx -> mx [weight = 0]
	fx -> fl [weight = 2]
	fl -> ml [weight = 2]
	mx -> ml [weight = 0]
}

ls -> tx [arrowhead = empty, weight = 0]
rl -> tx [arrowhead = empty, weight = 0]
pt -> fl [arrowhead = empty]
df -> mx [arrowhead = empty]
df -> ml [arrowhead = empty]

}")
```

</center>

In theory, the `character` class is considered enriched as a `term` class. In turn, a `term` class is considered enriched as a `formula` class. These enriched classes of <span style = "color:crimson">term</span> and <span style = "color:crimson">formula</span> could be implemented as vectors under this type of hierarchy, with emergent properties/attributes that develop in the hierarchy.

# Structure

```{r}
DiagrammeR::grViz("
digraph {

# Graph
graph [layout = dot, rankdir = RL]

# Nodes of object classes
term [shape = record, label='Term Record|+character|+side']

}")
```

